---
  output: github_document
indent: true
---

  <!-- README.md is generated from README.Rmd. Please edit that file -->

  ```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
  # out.width = "100%"
)
```
# CLIPreg

<!-- badges: start -->
  <!-- badges: end -->

  The goal of CLIPreg is to discover key RBP regulators in different datasets. It combines CLIP-seq with RNA- and RIBO-seq to calculate  enrichment of RBP and generate plots for publications.

## Installation

### Check and install required packages

Users may use following codes to check and install all the required packages.

``` r
list.of.packages <- c("ggplot2","pheatmap","grid","doParallel","foreach","data.table","fastmatch","GGally","ggnet","topGO","ALL","devtools")

## for package "ggplot2", "pheatmap", "grid", "doParallel", "foreach", "data.table", "fastmatch", "GGally"
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## for package "topGO", "ALL"
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages))  BiocManager::install(new.packages)
if("ggnet"%in%new.packages)  devtools::install_github("briatte/ggnet")
```

### Install CLIPreg

The source code of CLIPreg can be installed from [GitHub](https://github.com/) with:

  ``` r
devtools::install_github("SGDDNB/CLIPreg")
```


## Input datasets

CLIPreg requires 4 different inputs. A cluster input which is a dataframe containing geneID and the cluster given by DeltaTE output.
Curated CLIP-seq data that can be downloaded from . RIBO and RNA lfc and TPM. (TPM are optional)

```{r load library, message=FALSE}
## load libraries
library(CLIPreg)
```

Let's have a look at the cluster file

```{r load data, message=FALSE}
head(clusters)
```

```{r run enrichment analysis, message=FALSE}
# Make sure all your files are reachable from the same folder
folder=getwd()
RBP_data_ENCODE="rbp_gene_sel.txt"
RBP_data_POSTAR="rbp_gene_postar.txt"
clusters_file="Fibroblasts/Clusters.csv"
iterations=100000 # recommended for analysis

# The CLIPreg_V3 function requires a few minutes to run, save the data after running it to be sure not to lose it
res_Postar=CLIPReg_V3(folder=folder,RBP_data=RBP_data_POSTAR,cluster=clusters_file,iterations=100000)
res_Encode=CLIPReg_V3(folder=folder,RBP_data=RBP_data_ENCODE,cluster=clusters_file,iterations=100000)
save(res_Encode,file=paste0(folder,"/Fibroblasts/Res_RBP_Encode_V2.RData"))
save(res_Postar,file=paste0(folder,"/Fibroblasts/Res_RBP_Postar_V2.RData"))
load(paste0(folder,"/Fibroblasts/Res_RBP_Postar.RData"))
load(paste0(folder,"/Fibroblasts/Res_RBP_Encode_V2.RData"))
head(res_Encode[[1]])
```

The output from CLIPreg_V3 function is a list of dataframes named after the clusters. Each dataframe containing RBPs with their overlap details and pvalues.
Then we want to combine POSTAR and ENCODE to work with only one dataframe

```{r combining POSTAR and ENCODE, message=FALSE}

# Combine RBP results from Postar and Encode
all_RBPs=unique(c(res_Encode[[1]]$RBP,res_Postar[[1]]$RBP))

res_both=res_Postar
for (i in 1:nrow(res_Encode[[1]])) {
  r=res_Encode[[1]]$RBP[i]
  if (r%in%res_both[[1]]$RBP) {
    for (n in names(res_both)) {
      for (j in 2:6) {
        res_both[[n]][which(res_both[[1]]$RBP==r),j]=(res_both[[n]][which(res_both[[1]]$RBP==r),j]+res_Encode[[n]][i,j])/2
      }
    }
  } else {
    for (n in names(res_both)) {
      res_both[[n]]=rbind(res_both[[n]],res_Encode[[n]][i,])
    }
  }
}

```

We are only interested in the significant RBPs in each cluster.

```{r combining POSTAR and ENCODE, message=FALSE}
# keep only significant RBPs

to_keep=c()
for (n in names(res_both)) {
  to_keep=c(to_keep,res_both[[n]]$RBP[res_both[[n]]$pval<0.05])
}
to_keep=unique(to_keep)

for (n in names(res_both)) {
  res_both[[n]]=res_both[[n]][res_both[[n]]$RBP%in%to_keep,]
}
res=res_both
```

```{r loading LFC and TPM, message=FALSE}
# load fold change and tpm
rna_lfc = as.data.frame(read.delim(paste(sep="","Sig_genes_all3_RNA_time_series_pw_LRT.lfc")))
ribo_lfc = as.data.frame(read.delim(paste(sep="","Sig_genes_all3_Ribo_time_series_pw_LRT.lfc")))
ribOnly_lfc = as.data.frame(read.delim(paste(sep="","Sig_genes_all3_RibOnly_time_series_pw_LRT.lfc")))

tpm_rna = read.delim("tpm_rna.tab")
tpm_ribo = read.delim("tpm_ribo.tab ")

# IDENTIFIER and geneID loading
ensembl=read.csv("C:/Users/e0545037/Desktop/Baptiste/Bioinfo/Source/Homo_sapiens.GRCh38.95_ID_L_Name.csv")
ribo_lfc=ribo_lfc[rownames(ribo_lfc)%in%ensembl$geneID,]
index_geneID=match(rownames(ribo_lfc),ensembl$geneID)
ribo_lfc$IDENTIFIER=ensembl$IDENTIFIER[index_geneID]
```

Extract the RBP LFC from the RIBO_LFC

```{r Extract LFC of RBPs, message=FALSE}
# Change of RBPs
rbp_lfc=ribo_lfc[ribo_lfc$IDENTIFIER%in%res[[1]]$RBP,2:6]
rbp_lfc=rbp_lfc[!duplicated(rbp_lfc$IDENTIFIER),]
rownames(rbp_lfc)=rbp_lfc$IDENTIFIER
rbp_lfc=rbp_lfc[,-5]

for (n in names(res_both)) {
  res_both[[n]]=res_both[[n]][res_both[[n]]$RBP%in%ribo_lfc$IDENTIFIER,]
}

```

Generate and save heatmap to pdf. The heatmap represents the -logP of each RBP for each cluster. The blue RBPs are downregulated and the orange RBPs are upregulated. Only RBPs that are significant in at least one cluster are shown.

```{r Generate and save heatmap, message=FALSE}
# Heatmap of RBP scores

e=HeatmapRBP(res=res,RBP_change=rbp_lfc,grid=F)
e # Plot the heatmap

# Save the heatmap
location="C:/Users/e0545037/Desktop/Baptiste/PhD/CLIPreg/Fibroblasts/Heatmap_fibroblasts_V4.pdf"
n=length(e$tree_row$order)
pdf(location,8,3+n*0.15)
e
dev.off()
```

Simplify target list from Encode and Postar. The Targets list created here will be used for the plots coming further down.

```{r Simplify target list, message=FALSE}

# Get targets from POSTAR and ENCODE and combine them

TargetsEncode=getTarget(folder = folder,RBP_data = "rbp_gene_sel.txt",background = clusters$geneID)
TargetsPOSTAR=getTarget(folder = folder,RBP_data = "rbp_gene_postar.txt",background = clusters$geneID)
Targets=c(TargetsEncode,TargetsPOSTAR)
Targets=split(unlist(Targets, use.names = FALSE), rep(names(Targets), lengths(Targets)))
for (t in names(Targets)) {
  Targets[[t]]=unique(Targets[[t]])
}


```

We can generate a heatmap showing the LFC or TPM of one RBP's targets.

```{r Heatmap per target, message=FALSE}
# Heatmap of targets of 1 RBP

clusters=fread(paste0(folder,"/",clusters_file),header = T)
index_geneID_clusters=match(clusters$geneID,ensembl$geneID)
clusters$IDENTIFIER=ensembl$IDENTIFIER[index_geneID_clusters]

HeatmapTargets(folder="Folder",Clusters=clusters,RBP_name="YTHDF3",targets=Targets,
               counts_rna=counts_rna, counts_ribo=counts_ribo[,-16],scale=T,Not_DE=F)

HeatmapTargets(folder="Folder",Clusters=clusters,RBP_name="YTHDF3",targets=Targets,
               counts_rna=LFC_RNA[,c(2,8,14)], counts_ribo=LFC_RIBO[,c(3,8,13)],scale=F)
```

A bubble plot can be generated only if the clusters are the output of DeltaTE program.

```{r Bubble plot, message=FALSE}
# Bubble plot clusters if clusters are from DeltaTE
BubbleRBPs(RBP_res = res,clusters = clusters)
```

From the results, the user can choose a number of RBP to draw the network for by n. This will pick the n most changing RBPs.

```{r Network, message=FALSE}
# Draw network
Draw_network(rbp_lfc=rbp_lfc,res=res_both[-c(1,2)],clusters=clusters,n=5)
Draw_network_by_group(rbp_lfc=rbp_lfc,res=res_both[-c(1,2)],Targets=Targets,clusters=clusters,n=5)
```

```{r GO of specific nodes, message=FALSE}
# plot GO of nodes of size >th.
Plot_GO(rbp_lfc=rbp_lfc,res=res_both[-c(1,2)],Targets=Targets,clusters=clusters,n=5,
  all_genes=rownames(tpm_ribo),th=200,GO_to_show=3)
```

