---
  output: github_document
  indent: true
---

  <!-- README.md is generated from README.Rmd. Please edit that file -->

  ```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
  # out.width = "100%"
)
  ```
# CLIPreg

<!-- badges: start -->
  <!-- badges: end -->

  The goal of CLIPreg is to discover key RBP regulators in different datasets. It combines CLIP-seq with RNA- and RIBO-seq to calculate  enrichment of RBP and generate plots for publications.

## Installation

### Check and install required packages

Users may use following codes to check and install all the required packages.

``` r
list.of.packages <- c("ggplot2","grid","doParallel","foreach","data.table","fastmatch","GGally","ggnet","topGO","ALL","devtools","org.Hs.eg.db")

## for package "ggplot2", "pheatmap", "grid", "doParallel", "foreach", "data.table", "fastmatch", "GGally"
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

## for package "topGO", "ALL", "ggnet", "ComplexHeatmap","org.Hs.eg.db"
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages))  BiocManager::install(new.packages)
if("ggnet"%in%new.packages)  devtools::install_github("briatte/ggnet")
install_version("network", version = "1.16.1", repos = "http://cran.us.r-project.org")
if("ComplexHeatmap"%in%new.packages)  devtools::install_github("jokergoo/ComplexHeatmap")
```

### Install CLIPreg

The source code of CLIPreg can be installed from [GitHub](https://github.com/) with:

  ``` r
devtools::install_github("SGDDNB/CLIPreg")
```

CLIPreg requires 4 different inputs. A gene_groups input which is a dataframe containing geneID and the gene_groups given by DeltaTE output (ideally). The DeltaTE method can be found in the paper here: https://doi.org/10.1002/cpmb.108. It categorizes transcriptionally and/or translationally regulated genes into 4 categories: forwarded, intensified, exclusive and buffered, each category with an up or down direction.
Summarized CLIP-seq data, POSTAR and ENCODE are pre-loaded in the package. RIBO lfc and TPM.

```{r load library, message=FALSE,warning=FALSE}
## load libraries
library(CLIPreg)
library(ggplot2)
library(ComplexHeatmap)
library(grid)
library(doParallel)
library(ggnet)
library(topGO)
library(GGally)
library(data.table)
library(stringr)
```

### Basic usage

For testing the package with the example data
```{r load example data, message=FALSE, eval=F}
Example=Load_example()
```

For using your own data, you must specify a folder that contains at least 3 txt files that are named: gene_groups.txt, ribo_lfc.txt and ribo_tpm.txt. For the format of those 3 files, please refer to Advance usage step 1.

```{r load user data, message=FALSE, eval=F}
Input_data=Load_input_files(folder = "path/to/folder")
```

Run the analysis with all default parameters.
```{r Run default analysis, message=FALSE, eval=F}
results=run_CLIPreg(Example, is.example=T) # or run_CLIPreg(Input_data, is.example=F)
```

Generate the visual output from the results of the analysis. The Visualise function will create 2 new files in the folder given: an Rdata file containing the list object from the run_CLIPreg function and a pdf file with the main figures.

```{r RVisualize, message=FALSE,warning=FALSE, eval=F}
dir.create("Results_CLIPreg")
Visualise(results=results,folder="Results_CLIPreg")
```

### Advance usage

#### Step 1: Input datasets

Let's have a look at the gene_groups file from the example data. It consists of 2 columns containing the geneID and the gene groups for all the DE genes. This is optional, if you want to provide your own gene groups, go to next step.

```{r load gene_groups, message=FALSE}
data("gene_groups")
head(gene_groups)
```

You can input your own gene groups by using the function load_gene_groups() and giving the file location of your file as input.

```{r load gene_groups from user, message=FALSE}
#gene_groups=load_gene_groups(gene_groups_file = "path/to/gene_groups_file.txt")
```

Load POSTAR and ENCODE RBP data. Those are 2 public datasets which are processed in order to have lists of vector. Each vector is named after 1 RBP and contains the geneID of all the targets of that RBP. Combine both data in a target list. 

```{r load RBP data, message=FALSE}
data("RBP_ENCODE")
data("RBP_POSTAR")
Targets=combine_targets(RBP_list1=RBP_ENCODE,RBP_list2=RBP_POSTAR,background=gene_groups$geneID)
```

Load the fold change and identify the RBPs. If you have your own then provide your own data.

```{r loading LFC and TPM, message=FALSE}
# load fold change and tpm. optional if you want to use the input data
data("ribo_lfc")
data("tpm_ribo")

head(ribo_lfc)
#head(tpm_ribo)

# If you want to input your own data
#load_ribo_lfc(ribo_lfc_file = "ribo_lfc_file")
#load_ribo_tpm(ribo_tpm_file = "ribo_tpm_file")

```

#### Step 2: Data integration and analysis

Run the enrichment analysis using CLIPreg() function. This takes several minutes. If you want to have a look at the example results skip this step.

```{r run enrichment analysis, message=FALSE}
# The CLIPreg function requires a few minutes to run, save the data after running it to be sure not to lose it
# res_Postar=CLIPreg(RBP_data=RBP_POSTAR,gene_groups=gene_groups)
# res_Encode=CLIPreg(RBP_data=RBP_ENCODE,gene_groups=gene_groups)
# save(res_Encode,file="Res_RBP_Encode.RData")
# save(res_Postar,file="Res_RBP_Postar.RData")
```

If you want to get the results directly you can load it by using the example data results. The output of CLIPreg() is a list of dataframes. One dataframe per gene group containing the RBP and statistical information calculated during the analysis such as p-value and z-score.

```{r load res, message=FALSE}
data("res_Encode")
data("res_Postar")
head(res_Encode[[1]])
```

Then we want to combine POSTAR and ENCODE to work with only one dataframe and only keep RBPs that are significant in at least one gene group.

```{r combining POSTAR and ENCODE, message=FALSE}

res=CLIPreg::combine(res1=res_Encode,res2=res_Postar,FDR=0.1)

```

Extract the RBP LFC from the RIBO_LFC and keep only detected RBPs in res

```{r Extract LFC of RBPs, message=FALSE}
# Change of RBPs
rbp_lfc=rbp_change(res=res,ribo_lfc=ribo_lfc)

# Cure res data by removing RBPs that are not in the rbp_lfc dataframe
res=cure_res(res=res,rbp_lfc=rbp_lfc)
```


#### Step 3: Visualisation

Generate and save heatmap to pdf. The heatmap represents the -logP of each RBP for each gene group. The blue RBPs are downregulated and the orange RBPs are upregulated. Only RBPs that are significant in at least one gene group are shown.

```{r Generate and save heatmap,fig.height=15,fig.width=10}
# Heatmap of RBP scores

HeatmapRBP(res=res,rbp_lfc=rbp_lfc)
# Plot the heatmap with updated colors for the RBPs

# Save the heatmap
# e=HeatmapRBP(res=res,rbp_lfc=rbp_lfc)
# location="Heatmap_fibroblasts.pdf"
# n=length(e$tree_row$order)
# pdf(location,8,3+n*0.15)
# 
# dev.off()
```

A bubble plot can be generated only if the gene groups are the output of DeltaTE program.

```{r Bubble plot, message=FALSE}
# Bubble plot gene_groups if gene_groups are from DeltaTE. FDR has to be lower or equal to the FDR put in CLIPreg::combine() step
BubbleRBPs(res = res,gene_groups = gene_groups,rbp_lfc = rbp_lfc,FDR=0.1)
```

From the results, the user can choose a number of RBP to draw the network for by n. This will pick the n most changing RBPs.

```{r Network,message=FALSE,warning=FALSE}
# Draw network
Draw_network_by_group(rbp_lfc=rbp_lfc,res=res,Targets=Targets,gene_groups=gene_groups,n=5,forwarded = F)
```

```{r GO of specific nodes,fig.width=15, message=FALSE,warning=FALSE}
# plot GO
Plot_GO(rbp_lfc=rbp_lfc,res=res,Targets=Targets,gene_groups=gene_groups,n=5,
  tpm_ribo = tpm_ribo,th=200,GO_to_show=3,forwarded = F)
Plot_GO_node_name(rbp_lfc=rbp_lfc,res=res,Targets=Targets,gene_groups=gene_groups,n=5,
                  tpm_ribo = tpm_ribo,Nodes_to_keep=c(19,15),GO_to_show=3,forwarded = F)
Plot_GO_RBP(rbp_of_interest="QKI",tpm_ribo = tpm_ribo,Targets=Targets,gene_groups=gene_groups,GO_to_show=3)
```
